Biogeography & environmental conditions shape bacteriophage-bacteria networks across the human microbiome

### 2018-01-07 revfiles/ReviewerResponse.md

The authors developed a pipeline that predicts interactions between phages and their hosts from metagenomic data by exploiting similarities on the gene and protein level as well as CRISPRs with the random forest machine learning technique. They then compared the resulting phage-host networks across persons, time points and body sites. In general, this study is a valuable contribution, however some aspects of the network analysis are questionable, as detailed below.

## Major

I am suspicious of node removal studies. Even though the accuracy of the authors' inference procedure is much higher than those based on co-occurrence, the resulting networks still contain a number of false positives that can bias the node removal results. Even if the networks were perfect, where is the experimental evidence that an ecosystem will be less stable when separated into several components? Also, less stable in what sense? There are multiple definitions of ecological stability. Which one do the authors have in mind?

> **This is an important point and we agree with the reviewer that this warrants further clarification. Stability was not the best term here, but rather we think *robustness* better communicates our intent and interpretation. Previous work has shown that high connectance of an ecological network is associated with resistance to species extinction events and greater support of interactions through the network, such as horizontal gene transfer. For example, if a phage only has one host in a system, the removal of that host is expected to result in the extinction of the phage. If that phage had multiple hosts in a system, the removal of one host is less likely to result in extinction of that phage. We referred to this as "stability" in our manuscript, but we think that robustness is clearer. We went back and formally defined our use of "robustness" in the manuscript and updated the text to more clearly articulate this point. We had included citations around this relationship between connectance and robustness, and we added text to make this clearer in the manuscript as well. LINES 23, 33-34, 44, 65-66, 92-93, 220-223, 263, 317-318, 328-329, 337-344, 403-405**

@@

In the same context:
P. 14, l. 262: "These findings suggested that the community properties inferred from microbiome interaction network structures, such as stability (meaning a more highly connected network  is more stable because a randomly removed bacteria or phage node is less likely to divide or disintegrate [30,58] the overall network."
The authors define a network with many components after species removal as unstable and a network with a single component as stable. To avoid a tautology here, an external criterion of stability is needed. Assuming an appropriate external criterion is available (e.g. number of species that die upon removal of a species), is there any experimental evidence that species whose removal results in more components destabilize an ecosystem more than species whose removal results in fewer components?

> **Similar to above, we are talking more about robustness of the network to the random removal of nodes (phages and bacteria) instead of true ecological stability, which is a point we needed to make clearer. Previous work has shown that less highly connected networks are more likely to experience extinction events resulting from random microbe removal, as we discussed in the point above. Additionally, interactions such as horizontal gene transfer are more limited when there are less "connections" throughout the network. Previous work has supported these model assumptions and we cited some of those relevant pieces of work in the manuscript. LINES 23, 33-34, 44, 65-66, 92-93, 220-223, 263, 317-318, 328-329, 337-344, 403-405**

@@

I appreciate that the authors mention that no statement of significance can be made with the small sample numbers in the high-fat/low-fat and obese/healthy groups, but then why comparing network properties between these groups at all?

> **This analysis was meant to be descriptive and to primarily illustrate to the field what types of questions can be answered once we have more comprehensive datasets. We cannot be certain of the results because of the small sample size, but this illustrates the potential biology we can uncover with more comprehensive microbiome sampling and integrative analyses. LINES 230-231, 247-250**

@@

Lines 115-120: It is not clear at first sight that the paragraph is describing the bacterial fraction.

> **We agree and clarified this in the text.**

Lines 134-138: the description of the methods lacks some important description.
It is form example not clear if the nucleotide similarity feature is always computed as the best bitscore obtained on the training bacterial dataset, or if it uses another external bacterial dataset. The same remarks goes for the amino-acid similarity feature.

> **We agree that the description here is cursory, as we left the reader to see the methods section for more details. No external datasets were used, and we clarified this point in the text.**

### 2018-03-05 doc/manuscript.Rmd
Updated docs.

The bacterial and viral sequences from these studies were quality filtered and assembled into contigs. We further grouped the related bacterial and phage contigs into operationally defined units based on their k-mer frequencies and co-abundance patterns, similar to previous reports **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})** [@Roux:2016cc]. This was done both for dimensionality reduction and to prevent inflation of node counts by using contigs which are expected to represent multiple fragments from the same genomes. This was also done to create genome analogs that we could use in our classification model which was built using genome sequences. We referred to these operationally defined groups of related contigs as operational genomic units (OGUs). Each OGU represented a genomically similar sub-population of either bacteria or phages. Contig lengths within clusters ranged between $10^{3}$ and $10^{5.5}$ bp **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.

While supplementing the previous virome fraction quality control measures **(Table S1)** we found that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$). As most phages in these communities have been shown to be temperate (they integrate into bacterial genomes) using methods that included nucleotide alignments of phages to bacterial reference genomes, we interpreted this as confirmation that the majority of those phages were temperate and therefore shared elements with bacterial reference genomes -- a trend previously reported [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa]. We further confirmed that while the majority of these were expected to be temperate phages, there still remained a low level of bacterial sequence noise, which was evident as five (1.7%) OGUs with similar sequence elements to the bacterial 16S rRNA gene (blastn, e-value < $10^{-25}$, length > 1,000 bp). This is in line with previous work which has suggested that bacterial noise is an inevitable technical issue, and we considered this noise while interpreting our findings [@Roux:2013cr]. We also identified two OGUs as being complete, high confidence phages using the stringent Virsorter phage identification algorithm (class 1 confidence group) [@Roux:2015dh].

The whole metagenomic shotgun sequence samples, which primarily consisted of bacteria, had an average viral relative abundance of 0.4% **(Table S1)** [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei]. We found that only 2% (6 / 280 OGUs) of bacterial OGUs had significantly strong nucleotide similarity to phage reference genomes (e-value < $10^{-25}$) [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei]. No OGUs were confidently identified as lytic or temperate phage OGUs in the bacterial dataset using the Virsorter algorithm [@Roux:2015dh]. Together this suggests minimal bacterial OGU noise that should be considered in the study conclusions.

## Implementing Phage-Bacteria Interaction Prediction to Build a Community Network
We predicted which phage OGUs infected which bacterial OGUs using a random forest model trained on experimentally validated infectious relationships from six previous publications [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz]. Only bacteria and bacteriophages were used in the model. The training set contained 43 diverse bacterial species and 30 diverse phage strains, including both broad and specific ranges of infection **(Supplemental Figure \ref{ValidationOverview} A - B)**. While it is true that there are more known phages that infect bacteria, we were limited by the information confirming which phages do not infect certain bacteria and attempted to keep the numbers of positive and negative interactions similar. Phages with linear and circular genomes, as well as ssDNA and dsDNA genomes, were included in the analysis. Because we used DNA sequencing studies, RNA phages were not considered **(Supplemental Figure \ref{ValidationOverview} C-D)**. This training set included both positive relationships (a phage infects a bacterium) and negative relationships (a phage does not infect a bacterium). This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon previous work that only considered positive relationships [@Edwards:2015iz]. It is also worth noting that while a positive interaction is strong evidence that the interaction exists, we must also be conscious that negative interactions are only weak evidence for a lack of interaction because the finding could be the result of our inability to reproduce conditions in which those interactions occur. Altogether we decided to maintain a balanced dataset at the cost of under-sampling the available positive interaction information because the use of such a severely unbalanced dataset often results in over-fit and uninformative model training.

Four phage and bacterial genomic features were used in our random forest model to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq]. These features were calculated using the training set described above. We chose to utilize these metrics that directly compare nucleotide sequences between sample phages and bacteria, instead of relying on alignment to reference genomes or known marker genes, because we are extrapolating our model to highly diverse communities which we expect to diverge significantly from the available reference genomes. The resulting random forest model was assessed using nested cross validation, and the median area under its receiver operating characteristic (ROC) curve was `r signif(averageauc[c(averageauc$type %in% "AUC"), 2], digits = sigfig)`, the median model sensitivity was `r signif(averageauc[c(averageauc$type %in% "SENS"), 2], digits = sigfig)`, and median specificity was `r signif(averageauc[c(averageauc$type %in% "SPEC"), 2], digits = sigfig)` **(Figure \ref{RocCurve} A)**. This balance of confident true positives at the cost of fewer true negatives is ideal for this type of dataset which consists of primarily positive connections **(Supplemental Figure \ref{nestedperformance})**. Nested cross validation of the model demonstrated that the sensitivity and specificity of the model could vary but the overall model performance (AUC) remained more consistent **(Supplemental Figure \ref{crossval})**. This suggested that our model would perform with a similar overall accuracy despite changes in sensitivity/specificity trade-offs. The most important predictor in the model was amino acid similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**. Protein family interactions were moderately important to the model, and CRISPRs were largely uninformative, due to the minimal amount of identifiable CRISPRs in the dataset and their redundancy with the nucleotide similarity methods **(Figure \ref{RocCurve} B)**. Approximately one third of the training set relationships yielded no score and therefore were unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.

@@

We used our random forest model to classify the relationships between bacteria and phage operational genomic units, which were then used to build the interactive network. The master network, analogous to the universal microbiome network concept previously described [@Bashan:2016is], contained the three studies as sub-networks, which themselves each contained sub-networks for each sample **(Supplemental Figure \ref{NetworkDiagram})**. Metadata including study, sample ID, disease, and OGU abundance within the community were stored in the master network for parsing in downstream analyses **(Supplemental Figure \ref{NetworkDiagram})**. The phage and bacteria of the master network demonstrated both narrow broad ranges of infectious interactions **(Supplemental Figure \ref{inthm})**. Bacterial and phage relative abundance was recorded in each sample for each OGU and the weight of the edge connecting those OGUs was calculated as a function of those relative abundance values. The separate extraction of the phage and bacterial libraries ensured a more accurate measurement of the microbial communities, as has been outlined previously [@Kleiner:2015kd; @Meisel:2016iu]. The master network was highly connected and contained `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], big.mark = ",")` infectious relationships among `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], big.mark = ",")` nodes, representing `r prettyNum(nestats[c(nestats$cats %in% "PhageNodes"),"values"], big.mark = ",")` phages and `r prettyNum(nestats[c(nestats$cats %in% "BacteriaNodes"),"values"], big.mark = ",")` bacteria. Although the network was highly connected, not all relationships were present in all samples. Relationships were weighted by the relative abundances of their associated bacteria and phages. Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "skin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "skin"),"bacterianodes"], big.mark = ",")` bacteria, `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% total) and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"], big.mark = ",")` (`r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`%) of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**. Additionally, the subnetworks demonstrated narrow ranges of eccentricity across their nodes **(Supplemental Figure \ref{eccplot})**. The phages and bacteria in the diet and twin sample sets were more sparsely related, with the diet study consisting of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "diet"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "diet"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Edges"], big.mark = ",")` relationships, and the twin study containing `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "twin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "twin"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Edges"], big.mark = ",")` relationships **(Figure \ref{RocCurve} E - F)**. As an interesting validation measure, we identified five (1.7%) examples of phage OGUs which similar genomic elements to the four previously described, broadly infectious phages isolated from Lake Michigan (tblastx; e-value < $10^{-25}$) [@Malki:2015dw].

@@

@@

The average centrality of each sample was quantified using the weighted eigenvector centrality metric. Intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**. Also, skin sites that were occluded from the environment were much more highly connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.

@@

In addition to diet, the skin environment also influenced the microbiome interaction network structure. Network structure differed between environmentally exposed and occluded skin sites. The sites under greater environmental fluctuation and exposure (the exposed and intermittently exposed sites) were less connected and therefore were predicted to have a higher resilience against network degradation when random nodes are removed from the network. Likewise, intermittently moist sites demonstrated less connectedness than the moist and sebaceous sites. These findings agree with previous work that has shown that bacterial community networks differ by skin environment types [@Bashan:2016is]. Together these data suggested that body sites under greater degrees of fluctuation harbored less connected microbiomes that are potentially less robust to network disruption by extinction events. This points to a link between microbiome and environmental robustness toward network homeostasis and warrants further investigation.

While these findings take us an important step closer to understanding the microbiome through interspecies relationships, there are caveats to and considerations regarding our findings. First, as with most classification models, the infection classification model developed and applied is only as good as its training set -- in this case, the collection of experimentally-verified positive and negative infection data. Large-scale experimental screens for phage and bacteria infectious interactions that report high-confidence negative interactions (i.e., no infection) are desperately needed, as they would provide more robust model training and improved model performance. Furthermore, just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will improve future performance. Other creative and high performing models are currently being developed and the applications of these models to community network creation will continue to move this field forward [@Villarroel:2016jy; @Galiez:2017gm; @Ahlgren:2017gh].

@@

Although our models demonstrated satisfactory accuracy and overall performance, it was important to interpret our findings under the realization that our model was not perfect. This caveat is not new to the microbiome field, with a notable example being the use of 16S rRNA sequencing using the V4 variable region [@Meisel:2016iu]. Use of the V4 variable region excluded detection of major bacterial members, meaning that the findings were not able to completely describe the underlying biological environment. Despite this caveat, skin microbiome studies provided valuable biological insights by focusing on the community differences between groups (e.g. disease and healthy) which were both analyzed the same way. Similarly, here we focused our conclusions on the differences between the groups which were all treated the same, so that we can minimize our dependence on a perfect predictive model.

Finally, the networks in this study were built using operational genomic units (OGUs), which represented groups of highly similar bacteria or phage genomes or clustered genome fragments. Similar clustering definition and validation methods, both computational and experimental, have been implemented in other metagenomic sequencing studies, as well [@Minot:2012ed; @Deng:2014eb; @Brum:2015iaa; @Roux:2016cc]. These approaches could offer yet another level of sophistication to our network-based analyses. While this operationally defined clustering approach allows us to study whole community networks, our ability to make conclusions about interactions among specific phage or bacterial species or populations is inherently limited, compared to more focused, culture-based studies such as the work by Malki *et al* [@Malki:2015dw]. Future work must address this limitation, e.g., through improved binning methods and deeper metagenomic shotgun sequencing, but most importantly through an improved conceptual framing of what defines ecologically and evolutionarily cohesive units for both phage and bacteria [@Polz:2006fi]. Defining operational genomic units and their taxonomic underpinnings (e.g., whether OGU clusters represent genera or species) is an active area of work critical to the utility of this approach. As a first step, phylogenomic analyses have been performed to cluster cyanophage isolate genomes into informative groups using shared gene content, average nucleotide identity of shared genes, and pairwise differences between genomes [@Gregory:2016cg]. Such population-genetic assessment of phage evolution, coupled with the ecological implications of genome heterogeneity, will inform how to define nodes in future iterations of the ecological network developed here. Even though we are hesitant to speculate on phage host ranges at low taxonomic levels in our dataset, the data does agree with previous reports of instances of broad phage host range [@Malki:2015dw; @PaezEspino:2016gi]. Additionally, visualization of our dataset interactions using the heat map approach previously used in other host range studies, suggests a trend toward modular and nested tropism, but we do not have the strain-level resolution that powered those previous experimental studies.

@@

>>>>

The bacterial and viral sequences from these studies were quality filtered and assembled into contigs ("genomic fragments"). We further grouped the related bacterial and phage contigs into operationally defined units based on their k-mer frequencies and co-abundance patterns, similar to previous reports **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})** [@Roux:2016cc]. This was done both for dimensionality reduction and to prevent inflation of node counts by using contigs which are expected to represent multiple fragments from the same genomes. This was also done to create genome analogs that we could use in our classification model which was built using genome sequences. We referred to these operationally defined groups of related contigs as operational genomic units (OGUs). Each OGU represented a genomically similar sub-population of either bacteria or phages. Contig lengths within clusters ranged between $10^{3}$ and $10^{5.5}$ bp **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.

The original publications reported that the whole metagenomic shotgun sequence samples, which primarily consisted of bacteria, had an average viral relative abundance of 0.4% **(Table S1)** [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei]. We confirmed these reports by finding that only 2% (6 / 280 OGUs) of bacterial OGUs had significantly strong nucleotide similarity to phage reference genomes (e-value < $10^{-25}$) [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei]. Additionally, no OGUs were confidently identified as lytic or temperate phage OGUs in the bacterial dataset using the Virsorter algorithm [@Roux:2015dh].

We also supplemented the previous virome fraction quality control measures **(Table S1)** to find that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$). It is important to note that interpreting such alignment is complicated by the fact that most reference bacterial genomes also contain prophages (phages integrated into bacterial genomes), meaning we do not know to what extent the alignments were the result of bacterial contaminants in the virome fraction and what were true integrated prophages. As most phages in these communities have been shown to be temperate (they integrate into bacterial genomes) using methods that included nucleotide alignments of phages to bacterial reference genomes, we expected that a large fraction of those phages were temperate and therefore shared elements with bacterial reference genomes -- a trend previously reported [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa]. To ensure the purity of our sample sets, we supplemented the quality control measures by filtering out all OGUs that could be potential bacterial contaminants, as described previously [@Roux:2016cc]. This resulted in the removal of 143 OGUs that exhibited nucleotide similarity to bacterial genomes but no identifiable known phage elements. We were also able to identify two OGUs as representing **complete**, high confidence phages using the stringent Virsorter phage identification algorithm (class 1 confidence group) [@Roux:2015dh].

## Implementing Phage-Bacteria Interaction Prediction to Build a Community Network
We predicted which phage OGUs infected which bacterial OGUs using a random forest model trained on experimentally validated infectious relationships from six previous publications [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz]. Only bacteria and bacteriophages were used in the model. The training set contained 43 diverse bacterial species and 30 diverse phage strains, including both broad and specific ranges of infection **(Supplemental Figure \ref{ValidationOverview} A - B, Table S2)**. While it is true that there are more known phages that infect bacteria, we were limited by the information confirming which phages do not infect certain bacteria and attempted to keep the numbers of positive and negative interactions similar. Phages with linear and circular genomes, as well as ssDNA and dsDNA genomes, were included in the analysis. Because we used DNA sequencing studies, RNA phages were not considered **(Supplemental Figure \ref{ValidationOverview} C-D)**. This training set included both positive relationships (a phage infects a bacterium) and negative relationships (a phage does not infect a bacterium). This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon previous work that only considered positive relationships [@Edwards:2015iz]. It is also worth noting that while a positive interaction is strong evidence that the interaction exists, we must also be conscious that negative interactions are only weak evidence for a lack of interaction because the finding could be the result of our inability to reproduce conditions in which those interactions occur. Altogether we decided to maintain a balanced dataset at the cost of under-sampling the available positive interaction information because the use of such a severely unbalanced dataset often results in over-fit and uninformative model training. However, as an additional validation measure, we used the extensive additional positive interactions as a secondary dataset to confirm that we could identify infectious interactions from a more diverse bacterial and phage dataset. Using this approach, we confirmed that 382 additional phage reference genomes, representing a diverse range of phages, were matched to at least one reference bacterial host genome of the species that they were expected to infect **(Supplemental Figure \ref{secondval})**. Because the model was built on full genomes and used on OGUs, we also assessed whether our model was resilient to incomplete reference genomes. We found that the use of our model on random contigs representing as little as 50% length of the original reference phage and bacterial genomes resulted in minimal reduction in the ability of the model to identify relationships **(Supplemental Figure \ref{contigval})**.

Four phage and bacterial genomic features were used in our random forest model to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq]. These features were calculated using the training set described above. While the nucleotide and amino acid similarity metrics are expected to identify prophage signatures, the protein family interaction and CRISPR signatures are expected to aid in identifying lytic phages in addition to temperate phages. We chose to utilize these metrics that directly compare nucleotide sequences between sample phages and bacteria, instead of relying on alignment to reference genomes or known marker genes, because we are extrapolating our model to highly diverse communities which we expect to diverge significantly from the available reference genomes. The resulting random forest model was assessed using nested cross validation, and the median area under its receiver operating characteristic (ROC) curve was `r signif(averageauc[c(averageauc$type %in% "AUC"), 2], digits = sigfig)`, the median model sensitivity was `r signif(averageauc[c(averageauc$type %in% "SENS"), 2], digits = sigfig)`, and median specificity was `r signif(averageauc[c(averageauc$type %in% "SPEC"), 2], digits = sigfig)` **(Figure \ref{RocCurve} A)**. This balance of confident true positives at the cost of fewer true negatives is ideal for this type of dataset which consists of primarily positive connections **(Supplemental Figure \ref{nestedperformance})**. Nested cross validation of the model demonstrated that the sensitivity and specificity of the model could vary but the overall model performance (AUC) remained more consistent **(Supplemental Figure \ref{crossval})**. This suggested that our model would perform with a similar overall accuracy despite changes in sensitivity/specificity trade-offs. The most important predictor in the model was amino acid similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**. Protein family interactions were moderately important to the model, and CRISPRs were largely uninformative, due to the minimal amount of identifiable CRISPRs in the dataset and their redundancy with the nucleotide similarity methods **(Figure \ref{RocCurve} B)**. Approximately one third of the training set relationships yielded no score and therefore were unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.

@@

We used our random forest model to classify the relationships between bacteria and phage operational genomic units, which were then used to build the interactive network. The master network, analogous to the universal microbiome network concept previously described [@Bashan:2016is], contained the three studies as sub-networks, which themselves each contained sub-networks for each sample **(Supplemental Figure \ref{NetworkDiagram})**. Metadata including study, sample ID, disease, and OGU abundance within the community were stored in the master network for parsing in downstream analyses **(Supplemental Figure \ref{NetworkDiagram})**. The phage and bacteria of the master network demonstrated both narrow broad ranges of infectious interactions **(Supplemental Figure \ref{inthm})**. Bacterial and phage relative abundance was recorded in each sample for each OGU and the weight of the edge connecting those OGUs was calculated as a function of those relative abundance values. The separate extraction of the phage and bacterial libraries ensured a more accurate measurement of the microbial communities, as has been outlined previously [@Kleiner:2015kd; @Meisel:2016iu]. The master network was highly connected and contained `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], big.mark = ",")` infectious relationships among `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], big.mark = ",")` nodes, representing `r prettyNum(nestats[c(nestats$cats %in% "PhageNodes"),"values"], big.mark = ",")` phages and `r prettyNum(nestats[c(nestats$cats %in% "BacteriaNodes"),"values"], big.mark = ",")` bacteria. Although the network was highly connected, not all relationships were present in all samples. Relationships were weighted by the relative abundances of their associated bacteria and phages. Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "skin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "skin"),"bacterianodes"], big.mark = ",")` bacteria, `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% total) and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"], big.mark = ",")` (`r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`%) of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**. Additionally, the subnetworks demonstrated narrow ranges of eccentricity across their nodes **(Supplemental Figure \ref{eccplot})**. The phages and bacteria in the diet and twin sample sets were more sparsely related, with the diet study consisting of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "diet"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "diet"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Edges"], big.mark = ",")` relationships, and the twin study containing `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "twin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "twin"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Edges"], big.mark = ",")` relationships **(Figure \ref{RocCurve} E - F)**. As an interesting validation measure, we identified five (1.7%) examples of phage OGUs which contained similar genomic elements to the four previously described, broadly infectious phages isolated from Lake Michigan (tblastx; e-value < $10^{-25}$) [@Malki:2015dw].

@@

As an additional validation measure, we evaluated the tolerance of these findings to inaccuracies in the underlying network. As described above, our model is not perfect and there is likely to be noise from false positive and false negative predictions. We found that additional random noise, both by creating a fully connected graph or randomly reducing the number of edges to 60% of the original, changed the statistical significance values (p-values) of our findings but not by enough to change whether they were statistically significant (p-value < 0.05). Therefore the comparisons between groups are resilient to potential noise resulting from model false positive and false negative predictions **(Supplemental Figure \ref{idivnoise})**.

@@

The average centrality of each sample was quantified using the weighted eigenvector centrality metric. Intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly more connected than the moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**. Also, skin sites that were occluded from the environment were less connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**. We also confirmed that addition of noise to the underlying network, as described above, altered the values of statistical significance (p-values) but not by enough to change whether they were statistically significant **(Supplemental Figure \ref{skindivnoise})**.

@@

In addition to diet, the skin environment also influenced the microbiome interaction network structure. Network structure differed between environmentally exposed and occluded skin sites. The sites under greater environmental fluctuation and exposure (the exposed and intermittently exposed sites) were more connected and therefore were predicted to have a higher resilience against network degradation when random nodes are removed from the network. Likewise, intermittently moist sites demonstrated higher connectedness than the moist and sebaceous sites. These findings agree with previous work that has shown that bacterial community networks differ by skin environment types [@Bashan:2016is]. Together these data suggested that body sites under greater degrees of fluctuation harbored more highly connected microbiomes that are potentially more robust to network disruption by extinction events. This points to a link between microbiome and environmental robustness toward network homeostasis and warrants further investigation.

While these findings take us an important step closer to understanding the microbiome through interspecies relationships, there are caveats and considerations to our findings. First, as with most classification models, the infection classification model developed and applied is only as good as its training set -- in this case, the collection of experimentally-verified positive and negative infection data. Large-scale experimental screens for phage and bacteria infectious interactions that report high-confidence negative interactions (i.e., no infection) are desperately needed, as they would provide more robust model training and improved model performance. Furthermore, just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will improve future performance. Other creative and high performing models are currently being developed and the applications of these models to community network creation will continue to move this field forward [@Villarroel:2016jy; @Galiez:2017gm; @Ahlgren:2017gh].

@@

Although our models demonstrated satisfactory accuracy and overall performance, it was important to interpret our findings under the realization that our model was not perfect. This caveat is not new to the microbiome field, with a notable example being the use of 16S rRNA sequencing using the V4 variable region [@Meisel:2016iu]. Use of the V4 variable region excluded detection of major skin bacterial members, meaning that the findings were not able to completely describe the underlying biological environment. Despite this caveat, skin microbiome studies provided valuable biological insights by focusing on the community differences between groups (e.g. disease and healthy) which were both analyzed the same way. Similarly, here we focused our conclusions on the differences between the groups which were all treated the same, so that we can minimize our dependence on a perfect predictive model. We also provided explicit evidence that the introduction of noise equally to the compared groups did not significantly impact our findings.

Third, the networks in this study were built using operational genomic units (OGUs), which represented groups of highly similar bacteria or phage genomes or clustered genome fragments. Similar clustering definition and validation methods, both computational and experimental, have been implemented in other metagenomic sequencing studies, as well [@Minot:2012ed; @Deng:2014eb; @Brum:2015iaa; @Roux:2016cc]. These approaches could offer yet another level of sophistication to our network-based analyses. While this operationally defined clustering approach allows us to study whole community networks, our ability to make conclusions about interactions among specific phage or bacterial species or populations is inherently limited, compared to more focused, culture-based studies such as the work by Malki *et al* [@Malki:2015dw]. Future work must address this limitation, e.g., through improved binning methods and deeper metagenomic shotgun sequencing, but most importantly through an improved conceptual framing of what defines ecologically and evolutionarily cohesive units for both phage and bacteria [@Polz:2006fi]. Defining operational genomic units and their taxonomic underpinnings (e.g., whether OGU clusters represent genera or species) is an active area of work critical to the utility of this approach. As a first step, phylogenomic analyses have been performed to cluster cyanophage isolate genomes into informative groups using shared gene content, average nucleotide identity of shared genes, and pairwise differences between genomes [@Gregory:2016cg]. Such population-genetic assessment of phage evolution, coupled with the ecological implications of genome heterogeneity, will inform how to define nodes in future iterations of the ecological network developed here. Even though we are hesitant to speculate on phage host ranges at low taxonomic levels in our dataset, the data does agree with previous reports of instances of broad phage host range [@Malki:2015dw; @PaezEspino:2016gi]. Additionally, visualization of our dataset interactions using the heat map approach previously used in other host range studies, suggests a trend toward modular and nested tropism, but we do not have the strain-level resolution that powered those previous experimental studies.

Finally, it is important to note that our model was built using available full genomes with known interactions, while the experimental datasets resulted in OGUs created from metagenomic shotgun sequence sets, as described above. While this is an informative approach given available data, it is not ideal. We envision future work focusing on training models using metagenomic shotgun sample sets from "mock communities" of bacteria and phages with experimentally validated infectious relationships. This would also be more informative than relying on simulated metagenomic sample sets, whose use would result in models built on simulations and more assumptions instead of empirical data. Together this way the training set can be subjected to the same pre-processing, contig assembly, and OGU binning processes as the experimental data. Furthermore, exciting advances in long read sequencing platforms such as the Oxford Nanopore MinIon system will provide more accurate genomic scaffolds than *de novo* assembled contigs, allowing for more accurate training and predictions of our models. As discussed above, it is because our current model is susceptible to this noise that we focus our conclusions on comparisons between experimental groups that were both treated the same. This is also why it was important for us to evaluate the susceptibility of our results to noise caused by the less-than-perfect prediction model.

@@

Contig simulations were performed by randomly extracting a string of genomic nucleotides that represented a defined percent length of that genome. This was accomplished using our `RandomContigGenerator.pl`, which was published in the associated GitHub repository.
